---
title: "Methods 2 -- Portfolio Assignment 1"
output:
  pdf_document: default
---

- _Type:_ Group assignment
- _Due:_ 10 March 2024, 23:59

---

In the following exercises, you will be asked to generate and summarize simulations from statistical models. You should use what you have learned so far (i.e. for loop, if else statements, sampling from continuous and discrete distributions...) to generate observations and summarize your samples using (one of) the appropriate methods. You can find examples of how to do that in Ch. 5. Note that here we will only focus on generative models, several aspects for inference and hypothesis testing discussed in Ch. 4 are not directly needed in this context.

In the first exercise, we will assume that the population of interest has a proportion of 0.51 men and 0.49 women. Your model should reflect that.

Please submit your answers on GitHub Classroom.

---
```{r libs}
library(pacman)

pacman::p_load(tidyverse,
               ggpubr,
               ggplot2)


## to enable knitting to pdf

#install.packages('tinytex')
#tinytex::install_tinytex()


```

---

1. _(5.2 from ROS)_ __Continuous probability simulation:__ The logarithms of weights (in pounds) of men in the United States are approximately normally distributed with mean 5.13 and standard deviation 0.17; women’s log weights are approximately normally distributed with mean 4.96 and standard deviation 0.20. Suppose 10 adults selected at random step on an elevator with a capacity of 1750 pounds. What is the probability that their total weight exceeds this limit?

---

2. _(5.6 from ROS)_ __Propagation of uncertainty:__ We use a highly idealized setting to illustrate the use of simulations in combining uncertainties. Suppose a company changes its technology for widget production, and a study estimates the cost savings at \$5 per unit, but with a standard error of \$4. Furthermore, a forecast estimates the size of the market (that is, the number of widgets that will be sold) at 40 000, with a standard error of 10 000. Assuming these two sources of uncertainty are independent, use simulation to estimate the total amount of money saved by the new product (that is, savings per unit, multiplied by size of the market).

---

3. _(5.10 from ROS)_ __Inference for a ratio of parameters:__ A (hypothetical) study compares the costs and effectiveness of two different medical treatments.

    - In the first part of the study, the difference in costs between treatments A and B is estimated at \$600 per patient, with a standard error of \$400, based on a regression with 50 degrees of freedom.
    - In the second part of the study, the difference in effectiveness is estimated at 3.0 (on some relevant measure), with a standard error of 1.0, based on a regression with 100 degrees of freedom.
    - For simplicity, assume that the data from the two parts of the study were collected independently.

    Inference is desired for the incremental cost-effectiveness ratio: the difference between the average costs of the two treatments, divided by the difference between their average effectiveness, a problem discussed further by Heitjan, Moskowitz, and Whang (1999).

    (a) Create 1000 simulation draws of the cost difference and the effectiveness difference, and make a scatterplot of these draws.
    (b) Use simulation to come up with an estimate, 50% interval, and 95% interval for the incremental cost-effectiveness ratio.
    (c) Repeat, changing the standard error on the difference in effectiveness to 2.0.

---

a)
   first we made a function for the simulations, which then was called 1000 times via a for loop, then the cost and effectiveness differences are made into a dataframe and then plotted as a scatterplot

```{r thirdPart_Mie_a}

### a) 



## 1) create 1000 simulations draws of the cost difference and effectiveness difference by making a function

# declare variables
costDiff_sim = c()
effDiff_sim = c()


costDiff_se = 400
effDiff_se = 1

costDiff_mean = 600
effDiff_mean = 3



# function simulating the estimates of the cost and effectiveness differences
costDiff_simulator <- function()
{
  costDiff = rnorm(1, mean = costDiff_mean, sd = costDiff_se)
  
  effDiff = rnorm(1, mean = effDiff_mean, sd = effDiff_se)
  
  returnVector <- c(costDiff, effDiff)
  
  return(returnVector)
  
}

# for loop runing the function 1000 times and saves the estimates of the cost and effectiveness differences in a vector each
for (i in 1:1000)
  {
  simulations <- costDiff_simulator()
  costDiff_sim[i] <- simulations[1]
  effDiff_sim[i] <- simulations[2]
  }


# 2) creates a dataframe with the cost and effectiveness differences to use for plotting
data <- data.frame(
  x_costDiff = costDiff_sim, # the x values are the cost differences
  y_effDiff = effDiff_sim  # the y values are the effectiveness differences
)


# 3) makes simple scatterplot
plot(data$x_costDiff, data$y_effDiff, xlab="Average recent growth in personal income",
ylab="Incumbent party's vote share")




# makes a scatterplot
ggplot(data, aes(x = x_costDiff)) +
  geom_point(aes(y = y_effDiff), alpha = 0.6, color = 'red') +  # Plotting the data points
  labs(
    x = 'Differences in cost per patient',
    y = "Differences in effectiveness",
    title = 'Costs and effectiveness differences between 2 medical treatments',
  ) +
  theme_minimal()



```

b)
   firstly the estimate, the mean, of the incremental cost-effectiveness ratio is calculated. Then its se is calculated based on the relative standard error of the estimates for the cost and effectiveness differences, which is then used to get the 50% and 95% confidence interval.



```{r thirdPart_Mie_b}

### b) 


## 1) the incremental cost-effectiveness ratio is calculated based on the simulations from a)

ratioCostEff = costDiff_sim/effDiff_sim

#print(ratioCostEff)


## 2) the mean (the estimate) is found of the incremental cost-effectiveness ratio 

ratioCostEff_estimate = mean(ratioCostEff)



## 3) calculates the standard error of the estimate and uses it for making the 50% and 95% confidence interval for the incremental cost-effectiveness ratio


# finds the relative SE of the cost and the effectiveness differences

relativeSE_cost = costDiff_se/mean(costDiff_sim)

relativeSE_eff = effDiff_se/mean(effDiff_sim)



# calculates the se of the estimate
ratioCostEff_se = ratioCostEff_estimate * sqrt(relativeSE_cost^2+relativeSE_eff^2)




# calculates the 50% CI

startCI_50 = round( ratioCostEff_estimate - (2/3) * ratioCostEff_se, 2)

endCI_50 = round(ratioCostEff_estimate + (2/3) * ratioCostEff_se, 2)

CI_50 = print(paste("the 50% confidence interval is [", startCI_50, ", ", endCI_50, "]"))




# calculates the 95% CI

startCI_95 = round( ratioCostEff_estimate - 2 * ratioCostEff_se, 2)

endCI_95 = round( ratioCostEff_estimate + 2 * ratioCostEff_se, 2)

CI_95 = print(paste("the 95% confidence interval is [", startCI_95, ", ", endCI_95, "]"))



```


c)
   the standard error on the difference in effectiveness is changed from 1 to 2.
        then the code is as for a) and b)



```{r thirdPart_Mie_c}

### c) 

effDiff_se = 2  # sets the new standard error of the differences in effectiveness



## 1) the a) part of the code - making the simulations for the estimates for cost and effectiveness differences and plotting them as a scatterplot


# resets the vectors variables
costDiff_sim = c()
effDiff_sim = c()




# function simulating the estimates of the cost and effectiveness differences
costDiff_simulator <- function()
{
  costDiff = rnorm(1, mean = costDiff_mean, sd = costDiff_se)
  
  effDiff = rnorm(1, mean = effDiff_mean, sd = effDiff_se)
  
  returnVector <- c(costDiff, effDiff)
  
  return(returnVector)
  
}


# for loop runing the function 1000 times and saves the estimates of the cost and effectiveness differences in a vector each
for (i in 1:1000)
  {
  simulations <- costDiff_simulator()
  costDiff_sim[i] <- simulations[1]
  effDiff_sim[i] <- simulations[2]
  }


# creates a dataframe with the cost and effectiveness differences to use for plotting
data <- data.frame(
  x_costDiff = costDiff_sim, # the x values are the cost differences
  y_effDiff = effDiff_sim  # the y values are the effectiveness differences
)


# makes a scatterplot
ggplot(data, aes(x = x_costDiff)) +
  geom_point(aes(y = y_effDiff), alpha = 0.6, color = 'red') +  # Plotting the data points
  labs(
    x = 'Differences in cost per patient',
    y = "Differences in effectiveness",
    title = 'Costs and effectiveness differences between 2 medical treatments',
  ) +
  theme_minimal()






## 2) the b) part of the code - calculating an estimate of the incremental cost-effectiveness ratio. Then the standard error used for the 50% and 95% confidence interval


## the incremental cost-effectiveness ratio is calculated based on the simulations from a)
ratioCostEff = costDiff_sim/effDiff_sim


## the mean (the estimate) is found of the incremental cost-effectiveness ratio 
ratioCostEff_estimate = mean(ratioCostEff)



## calculates the standard error of the estimate and uses it for making the 50% and 95% confidence interval for the incremental cost-effectiveness ratio


# finds the relative SE of the cost and the effectiveness differences

relativeSE_cost = costDiff_se/mean(costDiff_sim)

relativeSE_eff = effDiff_se/mean(effDiff_sim)



# calculates the se of the estimate
ratioCostEff_se = ratioCostEff_estimate * sqrt(relativeSE_cost^2+relativeSE_eff^2)




# calculates the 50% CI

startCI_50 = round( ratioCostEff_estimate - (2/3) * ratioCostEff_se, 2)

endCI_50 = round(ratioCostEff_estimate + (2/3) * ratioCostEff_se, 2)

CI_50 = print(paste("the 50% confidence interval is [", startCI_50, ", ", endCI_50, "]"))




# calculates the 95% CI

startCI_95 = round( ratioCostEff_estimate - 2 * ratioCostEff_se, 2)

endCI_95 = round( ratioCostEff_estimate + 2 * ratioCostEff_se, 2)

CI_95 = print(paste("the 95% confidence interval is [", startCI_95, ", ", endCI_95, "]"))



```


